---
title: "Hands-on Exercise 05.3"
subtitle: "Creating Parallel Coordinates Plots with R"
---

## **Overview**

Parallel coordinates plot is a data visualisation specially designed for visualising and analysing multivariate, numerical data. It is ideal for comparing multiple variables together and seeing the relationships between them. For example, the variables contribute to Happiness Index. Parallel coordinates was invented by [Alfred Inselberg](http://www.math.tau.ac.il/~aiisreal/) in the 1970s as a way to visualize high-dimensional data. This data visualisation technique is more often found in academic and scientific communities than in business and consumer data visualizations. As pointed out by [Stephen Few](https://www.perceptualedge.com/articles/b-eye/parallel_coordinates.pdf)(2006), \"This certainly isn\'t a chart that you would present to the board of directors or place on your Web site for the general public. In fact, the strength of parallel coordinates isn\'t in their ability to communicate some truth in the data to others, but rather in their ability to bring meaningful multivariate patterns and comparisons to light when used interactively for analysis.\" For example, parallel coordinates plot can be used to characterise clusters detected during customer segmentation.

By the end of this hands-on exercise, you will gain hands-on experience on:

-   plotting statistic parallel coordinates plots by using *ggparcoord()* of **GGally** package,

-   plotting interactive parallel coordinates plots by using **parcoords** package, and

-   plotting interactive parallel coordinates plots by using **parallelPlot** package.

### **Installing and Launching R Packages**

For this exercise, the **GGally**, **parcoords**, **parallelPlot** and **tidyverse** packages will be used.

The code chunks below are used to install and load the packages in R.

```{r}
pacman::p_load(GGally, parallelPlot, tidyverse)
```

### **Data Preparation**

In this hands-on exercise, the World Happinees 2018 (http://worldhappiness.report/ed/2018/) data will be used. The data set is download at https://s3.amazonaws.com/happiness-report/2018/WHR2018Chapter2OnlineData.xls. The original data set is in Microsoft Excel format. It has been extracted and saved in csv file called **WHData-2018.csv**.

In the code chunk below, `read_csv()` of **readr** package is used to import *WHData-2018.csv* into R and save it into a tibble data frame object called *wh*.

```{r}
wh <- read_csv("data/WHData-2018.csv")
```
