---
title: "Data Visualization Makeover 04"
subtitle: "Putting Visual Analytics into Practical Use"
author: "Bryant Philippe Lee"
date: February 26, 2022
---

## Objective

The objective is to uncover the impact of COVID-19 as well as the global economic and political dynamic in 2022 on Singapore bi-lateral trade (i.e. Import, Export and Trade Balance) by using appropriate analytical visualisation techniques to enhance user and data discovery experiences.

The visualization is created using the **Merchandise Trade** provided by **Department of Statistics, Singapore (DOS)*.***\

## Getting Started - Data Loading and Processing

### Installing and loading the required libraries

```{r}
pacman::p_load(scales, viridis, lubridate, ggthemes, gridExtra, tidyverse, readxl, knitr, data.table, readxl, lubridate, stringr, dplyr)
```

### Importing Data

This code chunk is to import the data from Merchandise Trade Excel File. The following sheets are within the file:

-   T1 = Import Data

-   T2 = Export Data

The data shows the import export relationship between Singapore and the other nations in terms of Million Dollar or Thousand Dollar

::: callout-important
#### Data Structure

The Merchandise Trade data set is labeled with Thousand Dollars and Million Dollars. Further examination shows the following:

**Million Dollars - Continent:**

-   Listed under this category are continental data and values, as such are the summation of nations within these continents.

-   The summation of this continents (except for European Union) is = Total Merchandise Exports

**Thousand Dollars - Nations:**

-   Nation specific data and values
:::

```{r}
import_continent <- read_excel("data/Merchandise_Trade.xlsx", 
                          sheet = "T1",
                          range = cell_rows(10:16))%>%
  select(`Data Series`, contains(c("2019","2020", "2021", "2022"))) %>%
  slice(-1) %>%
  set_names(if_else(names(.) == "Data Series", "Country", names(.)))

import_continent$Country <- str_replace_all(import_continent$Country, "\\s*\\(.*?\\)", "")

import_continent
```

```{r}
export_continent <- read_excel("data/Merchandise_Trade.xlsx", 
                          sheet = "T2",
                          range = cell_rows(10:16))%>%
  select(`Data Series`, contains(c("2019","2020", "2021", "2022"))) %>%
  slice(-1) %>%
  set_names(if_else(names(.) == "Data Series", "Country", names(.)))

export_continent$Country <- str_replace_all(export_continent$Country, "\\s*\\(.*?\\)", "")

export_continent
```

```{r}
import_nation <- read_excel("data/Merchandise_Trade.xlsx", 
                          sheet = "T1",
                          range = cell_rows(10:129))%>%
  select(`Data Series`, contains(c("2019","2020", "2021", "2022"))) %>%
  slice(-c(1:7))%>%
  set_names(if_else(names(.) == "Data Series", "Country", names(.)))

import_nation$Country <- str_replace_all(import_nation$Country, "\\s*\\(.*?\\)", "")

import_nation
```

```{r}
export_nation <- read_excel("data/Merchandise_Trade.xlsx", 
                          sheet = "T2",
                          range = cell_rows(10:101))%>%
  select(`Data Series`, contains(c("2019","2020", "2021", "2022"))) %>%
  slice(-c(1:7)) %>%
  set_names(if_else(names(.) == "Data Series", "Country", names(.)))

export_nation$Country <- str_replace_all(export_nation$Country, "\\s*\\(.*?\\)", "")

export_nation
```

### Exploring and Cleaning the Data

#### Pivoting Dataset

This section ensures that the datetime column labels are represented correctly. The dataset must be pivoted to reflect that.

```{r}
imports_C_pivot <- import_continent %>%
  pivot_longer(
    cols = !"Country",
    names_to = "Month-Year",
    values_to = "Amount",
    values_drop_na = TRUE
  )

imports_C_pivot
```

```{r}
imports_N_pivot <- import_nation %>%
  pivot_longer(
    cols = !"Country",
    names_to = "Month-Year",
    values_to = "Amount",
    values_drop_na = TRUE
  )

imports_N_pivot
```

```{r}
exports_C_pivot <- export_continent %>%
  pivot_longer(
    cols = !"Country",
    names_to = "Month-Year",
    values_to = "Amount",
    values_drop_na = TRUE
  )

exports_C_pivot
```

```{r}
exports_N_pivot <- export_nation %>%
  pivot_longer(
    cols = !"Country",
    names_to = "Month-Year",
    values_to = "Amount",
    values_drop_na = TRUE
  )

exports_N_pivot
```

#### Organizing Month-Year

This Section converts the "Month-Year" column to datetime.

```{r}
imports_C_pivot$`Month-Year` <- as.Date(paste(imports_C_pivot$`Month-Year`, "01"), format = "%Y %b %d")

imports_C_pivot
```

```{r}
imports_N_pivot$`Month-Year` <- as.Date(paste(imports_N_pivot$`Month-Year`, "01"), format = "%Y %b %d")

imports_N_pivot
```

```{r}
exports_C_pivot$`Month-Year` <- as.Date(paste(exports_C_pivot$`Month-Year`, "01"), format = "%Y %b %d")

exports_C_pivot
```

```{r}
exports_N_pivot$`Month-Year` <- as.Date(paste(exports_N_pivot$`Month-Year`, "01"), format = "%Y %b %d")

exports_N_pivot
```

#### Merge Dataset for both Continental and National Values

This section creates the consolidated dataset for both continental and national datasets with the following improvements:

-   Net Trade = Exports - Imports

-   Omit of all NA values

```{r}
continent <- merge(x=imports_C_pivot , y=exports_C_pivot, by= c("Country", "Month-Year"), all.x = TRUE, all.y = TRUE)

Net_Trade = continent$Amount.y - continent$Amount.x

continent$Net_Trade <- Net_Trade

na.omit(continent) %>%
  rename(
    Continent = Country,
    Date = `Month-Year`,
    Imports = Amount.x,
    Exports = Amount.y
  )
```

```{r}
nation <- merge(x=imports_N_pivot , y=exports_N_pivot, by= c("Country", "Month-Year"), all.x = TRUE, all.y = TRUE)

Net_Trade = nation$Amount.y - nation$Amount.x

nation$Net_Trade <- Net_Trade

na.omit(nation) %>%
  rename(
    Date = `Month-Year`,
    Imports = Amount.x,
    Exports = Amount.y
  )
```

## Continental Trade Analysis - Visual Exploration

This section explores macro trade relationship of Singapore on the continental level.

```{r}
#| fig-height: 12
#| fig-width: 12

ggplot(continent,
       aes(x = continent$`Month-Year`, y = continent$Amount.x, color = continent$Country)) +
  geom_line()+
  ggtitle("Import in Million Dollars by Continent") +
  xlab("Date") +
  ylab("Import in Million Dollars") +
  scale_x_date(date_labels = "%b-%y")
```
