---
title: "Data Visualization Makeover 03"
subtitle: "Putting Visual Analytics into Practical Use"
author: "Bryant Philippe Lee"
date: February 15, 2022
---

## Objective

The objective of this exercise is to uncover the salient patterns of the resale prices of public housing property by residential towns and estates in Singapore by using appropriate analytical visualisation techniques and apply appropriate interactive techniques to enhance user and data discovery experiences.

The visualization is created using the ***Resale flat princes based on registration date from Jan-2017 onwards ([Data.gov.sg](https://data.gov.sg/)).***

### Scope and Limitation

For the purpose of this study, the focus should be on **3-ROOM, 4-ROOM and 5-ROOM types**. You can choose to focus on either one housing type or multiple housing types. The study period should be on **2022**.

## Getting Started - Data Loading and Processing

### Installing and loading the required libraries

Before we get started, it is important for us to ensure that the required R packages have been installed.

```{r}
pacman::p_load(tidyverse, ggplot2, knitr, plotly, skimr, questionr, funModeling, sf, tmap, quanteda, NLP, ggiraph, ggstatsplot, dplyr, crosstalk, DT, ggdist, gganimate, ggpubr)
```

### Importing Data

This code chunk is to import the data from resale-flat-prices-based-on-registration-date-from-jan-2017-onwards.csv file to the Quarto/R page.

```{r}
total_flat_data <- read_csv("data/resale-flat-prices-based-on-registration-date-from-jan-2017-onwards.csv")
```

### Filtering the Dataset

This section deals with limiting the data to the following criteria:

-   The year should be **2022**

-   The flat_type should be only **3-ROOM, 4-ROOM and 5-ROOM**

::: callout-note
This code uses the dplyr package to filter out the rows with year 2022 and correct flat type
:::

```{r}
filter_flat_data <- filter(total_flat_data, grepl('2022', month) & flat_type %in% c("3 ROOM", "4 ROOM","5 ROOM"))

filter_flat_data
```

### Exploring the Data

This section of the code aims to ensure that the scope and limitation are followed and the data are listed correctly

```{r}
skimr::skim(filter_flat_data)
```

```{r}
glimpse(filter_flat_data, 60)
```

```{r}
unique(filter_flat_data$month)
```

```{r}
unique(filter_flat_data$flat_type)
```

```{r}
unique(filter_flat_data$storey_range)
```

::: callout-note
\*storey_range is listed as a string and needed for the arrangement to be corrected moving foreward
:::

```{r}
unique(filter_flat_data$floor_area_sqm)
```

```{r}
unique(filter_flat_data$remaining_lease)
```

::: callout-note
\*remaining lease is listed as a string and needed for the arrangement to be corrected moving foreward
:::

```{r}
summary(filter_flat_data$resale_price)
```

### Correcting the Data

This section deals with correcting the data last inspected previously.

#### Correcting the Storey Arrangement

The code adds corrects the order of the storey arrangement

```{r}
storey_correct <- c("01 TO 03", "04 TO 06", "07 TO 09", "10 TO 12", "13 TO 15", "16 TO 18", "19 TO 21", "22 TO 24", "25 TO 27", "28 TO 30", "31 TO 33", "34 TO 36", "37 TO 39", "40 TO 42", "43 TO 45", "46 TO 48", "49 TO 51")  

use_flat_data <- filter_flat_data %>%
  mutate (storey_range = factor(storey_range, levels = storey_correct)) %>%
  ungroup()
```

```{r}
ggplot(data = use_flat_data,
       aes(y = storey_range)) +
  geom_bar() 
```

#### Converting Remaining Lease from String to Num

The code adds a new integer column derived from the remaining_lease string column.

::: callout-note
This code uses the gsub-regex function to get the integer within the string.
:::

```{r}
lease_ryear <- as.numeric(gsub("([0-9]+).*$", "\\1", filter_flat_data$remaining_lease))

use_flat_data$lease_ryear <- lease_ryear

use_flat_data
```

## Data Exploration through Statistical Visualization

This section presents the statistical visualization and explanation of the data.

### One-way ANOVA Test between Flat Type and Resale Price

```{r}
#| fig-height: 12
#| fig-width: 12
#| code-fold: true

ggbetweenstats(
  data = use_flat_data,
  x = flat_type,
  y = resale_price, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```
