---
title: "Data Visualization Makeover 04"
subtitle: "Putting Visual Analytics into Practical Use"
author: "Bryant Philippe Lee"
date: February 26, 2022
---

## Objective

The objective is to uncover the impact of COVID-19 as well as the global economic and political dynamic in 2022 on Singapore bi-lateral trade (i.e. Import, Export and Trade Balance) by using appropriate analytical visualisation techniques to enhance user and data discovery experiences.

The visualization is created using the **Merchandise Trade** provided by **Department of Statistics, Singapore (DOS)*.***\

## Getting Started - Data Loading and Processing

### Installing and loading the required libraries

```{r}
pacman::p_load(scales, viridis, lubridate, ggthemes, gridExtra, tidyverse, readxl, knitr, data.table, readxl, lubridate, stringr, dplyr, gganimate, transformr, patchwork, countrycode)
```

### Importing Data

This code chunk is to import the data from Merchandise Trade Excel File. The following sheets are within the file:

-   T1 = Import Data

-   T2 = Export Data

The data shows the import export relationship between Singapore and the other nations in terms of Million Dollar or Thousand Dollar

::: callout-important
#### Data Structure

The Merchandise Trade data set is labeled with Thousand Dollars and Million Dollars. Further examination shows the following:

**Million Dollars - Continent:**

-   Listed under this category are continental data and values, as such are the summation of nations within these continents.

-   The summation of this continents (except for European Union as it is part of Europe) is = Total Merchandise Exports

**Thousand Dollars - Nations:**

-   Nation specific data and values
:::

```{r}
import_continent <- read_excel("data/Merchandise_Trade.xlsx", 
                          sheet = "T1",
                          range = cell_rows(10:16))%>%
  select(`Data Series`, contains(c("2019","2020", "2021", "2022"))) %>%
  slice(-1) %>%
  set_names(if_else(names(.) == "Data Series", "Country", names(.)))

import_continent$Country <- str_replace_all(import_continent$Country, "\\s*\\(.*?\\)", "")

import_continent
```

```{r}
export_continent <- read_excel("data/Merchandise_Trade.xlsx", 
                          sheet = "T2",
                          range = cell_rows(10:16))%>%
  select(`Data Series`, contains(c("2019","2020", "2021", "2022"))) %>%
  slice(-1) %>%
  set_names(if_else(names(.) == "Data Series", "Country", names(.)))

export_continent$Country <- str_replace_all(export_continent$Country, "\\s*\\(.*?\\)", "")

export_continent
```

```{r}
import_nation <- read_excel("data/Merchandise_Trade.xlsx", 
                          sheet = "T1",
                          range = cell_rows(10:129))%>%
  select(`Data Series`, contains(c("2019","2020", "2021", "2022"))) %>%
  slice(-c(1:7))%>%
  set_names(if_else(names(.) == "Data Series", "Country", names(.)))

import_nation$Country <- str_replace_all(import_nation$Country, "\\s*\\(.*?\\)", "")

import_nation
```

```{r}
export_nation <- read_excel("data/Merchandise_Trade.xlsx", 
                          sheet = "T2",
                          range = cell_rows(10:101))%>%
  select(`Data Series`, contains(c("2019","2020", "2021", "2022"))) %>%
  slice(-c(1:7)) %>%
  set_names(if_else(names(.) == "Data Series", "Country", names(.)))

export_nation$Country <- str_replace_all(export_nation$Country, "\\s*\\(.*?\\)", "")

export_nation
```

### Exploring and Cleaning the Data

#### Pivoting Dataset

This section ensures that the datetime column labels are represented correctly. The dataset must be pivoted to reflect that.

```{r}
imports_C_pivot <- import_continent %>%
  pivot_longer(
    cols = !"Country",
    names_to = "Month-Year",
    values_to = "Amount",
    values_drop_na = TRUE
  )

imports_C_pivot
```

```{r}
imports_N_pivot <- import_nation %>%
  pivot_longer(
    cols = !"Country",
    names_to = "Month-Year",
    values_to = "Amount",
    values_drop_na = TRUE
  )

imports_N_pivot
```

```{r}
exports_C_pivot <- export_continent %>%
  pivot_longer(
    cols = !"Country",
    names_to = "Month-Year",
    values_to = "Amount",
    values_drop_na = TRUE
  )

exports_C_pivot
```

```{r}
exports_N_pivot <- export_nation %>%
  pivot_longer(
    cols = !"Country",
    names_to = "Month-Year",
    values_to = "Amount",
    values_drop_na = TRUE
  )

exports_N_pivot
```

#### Organizing Month-Year

This Section converts the "Month-Year" column to datetime.

```{r}
imports_C_pivot$`Month-Year` <- as.Date(paste(imports_C_pivot$`Month-Year`, "01"), format = "%Y %b %d")

imports_C_pivot
```

```{r}
imports_N_pivot$`Month-Year` <- as.Date(paste(imports_N_pivot$`Month-Year`, "01"), format = "%Y %b %d")

imports_N_pivot
```

```{r}
exports_C_pivot$`Month-Year` <- as.Date(paste(exports_C_pivot$`Month-Year`, "01"), format = "%Y %b %d")

exports_C_pivot
```

```{r}
exports_N_pivot$`Month-Year` <- as.Date(paste(exports_N_pivot$`Month-Year`, "01"), format = "%Y %b %d")

exports_N_pivot
```

#### Merge Dataset for both Continental and National Values

This section creates the consolidated dataset for both continental and national datasets with the following improvements:

-   Net Trade = Exports - Imports

-   Omit of all NA values

```{r, results='hide'}
continent <- merge(x=imports_C_pivot , y=exports_C_pivot, by= c("Country", "Month-Year"), all.x = TRUE, all.y = TRUE)

Net_Trade = continent$Amount.y - continent$Amount.x

continent$Net_Trade <- Net_Trade

na.omit(continent) %>%
  rename(
    Continent = Country,
    Date = `Month-Year`,
    Imports = Amount.x,
    Exports = Amount.y
  )
```

```{r, results='hide'}
nation <- merge(x=imports_N_pivot , y=exports_N_pivot, by= c("Country", "Month-Year"), all.x = TRUE, all.y = TRUE)

Net_Trade = nation$Amount.y - nation$Amount.x

nation$Net_Trade <- Net_Trade

na.omit(nation) %>%
  rename(
    Date = `Month-Year`,
    Imports = Amount.x,
    Exports = Amount.y
  )
```

## Continental Trade Analysis - Visual Exploration

This section explores macro trade relationship of Singapore on the continental level.

### Line Graph Visualization for Imports and Exports

```{r}
#| fig-height: 16
#| fig-width: 14
#| code-fold: true

continent_import <- ggplot(continent,
       aes(x = `Month-Year`, y = Amount.x, color = Country)) +
  geom_line()+
  labs (x = "Date",
        y = "Import in Million Dollars",
        fill = "Continent",
        title = "Import in Million Dollars by Continent") +
  scale_x_date(date_labels = "%b-%y")+
  theme_bw() 

continent_export <- ggplot(continent,
       aes(x = `Month-Year`, y = Amount.y, color = Country)) +
  geom_line()+
  labs (x = "Date",
        y = "Export in Million Dollars",
        fill = "Continent",
        title = "Export in Million Dollars by Continent") +
  scale_x_date(date_labels = "%b-%y")+
  theme_bw() 

patchwork <- (continent_import / continent_export)
patchwork
```

A simple beginning in visual analysis is a line chart to see the time series impact on both imports and exports. Here is the following insights:

-   **Asia remains as the highest import export partner for Singapore.** Consistently, from 2019 to 2022, Asia remains as the top import export partner of Singapore due to proximity. America and Europe follows as close 2nd and 3rd.

-   **COVID 19 had none to minimal impact to Singapore's Macro Trade Relationship.** Year 2019 and 2020 - 2021 show no distinct drop in import and export amount (drop = significant change in numbers as minimal up and downs are expected).

### Animated Time Series Scatter Plot

```{r}
#| fig-height: 16
#| fig-width: 14
#| code-fold: true

ggplot(continent, aes(x = Amount.x, y = Amount.y, size = Net_Trade, color = Country)) + 
  geom_point(show.legend = FALSE, alpha = 0.7) +
  scale_color_viridis_d() +
  scale_size(range = c(2, 12)) +
  transition_time(continent$`Month-Year`) +
  labs(title = "Animated Scatterplot per Continent per Year: {frame_time}", x = "Imports", y = "Exports", fill = "Continent") +
  theme(legend.position = "bottom") + 
  theme_bw() + 
  facet_wrap(~Country)
```

The Animated Scatter plot refocuses that Asia is the highest import export trade partner for Singapore, though there is an additional insight

-   Asia remains as the highest import export partner and moving to 2022, both import and export increases as it reaches 2022 as the other continent remained constant.

-   Europe shows periods of fluctuations in the index (export/import) as compared to the more stable continents.

### Next Step

-   Visualize and Analyze nations within Asia and Europe separately to check and describe the time series progress of the import export trade amount.

## Asian Trade Analysis - Visual Exploration

This section explores the macro economic trade relation of Singapore to its Asian Trade partners.

### Identifying Country - Continent Labels

This section adds a continent label to the nation database identifying which continent a country belongs to.

::: callout-note
#### countrycode() Package

You can use the countrycode() package to identify the continent and region of the country

-   **continent** = Continent as defined in the World Bank Development Indicators

-   **un.regionsub.name**: United Nations sub-region name
:::

```{r, results='hide'}
nation$Continent <- countrycode(sourcevar = nation$Country,
                            origin = "country.name",
                            destination = "continent")

nation$Continent[nation$Country == "Other Country"] <- "Unknown"

nation$Region <- countrycode(sourcevar = nation$Country,
                            origin = "country.name",
                            destination = "un.regionsub.name")

nation$Region[nation$Country == "Other Country"] <- "Unknown"

nation <- na.omit(nation)

nation
```

```{r, results='hide'}
nation_asia <- nation %>%
  filter(str_detect(Continent, "Asia")) %>%
  select(-(Continent))

nation_asia
```

```{r, results='hide'}
nation_eu <- nation %>%
  filter(str_detect(Continent, "Europe")) %>%
  select(-(Continent))

nation_eu
```

### Asia Sub Region Visualization

```{r}
#| fig-height: 16
#| fig-width: 14
#| code-fold: true

ggplot(nation_asia, aes(x = Amount.x, y = Amount.y, size = Net_Trade, color = Country)) + 
  geom_point(show.legend = FALSE, alpha = 0.7) +
  scale_color_viridis_d() +
  scale_size(range = c(2, 12)) +
  transition_time(nation_asia$`Month-Year`) +
  labs(title = "Animated Scatterplot per Asian Region per Year: {frame_time}", x = "Imports", y = "Exports", fill = "Region") +
  theme(legend.position = "bottom") + 
  theme_bw() + 
  facet_wrap(~Region)
```

Based on the Animated Scatter Plot, the following insights are:

-   East Asia shows high import and export contribution. It is noted that there is a nation that have seen fluctuations in positioning in 2021 - 2022

-   Southeast Asia shows a more stable amount, thoigh there is a nation with more fluctuating position

### East Asian Import Export Visualization

This section shows the trade relation in East Asia.

```{r}
#| fig-height: 12
#| fig-width: 14
#| code-fold: true

nation_asia_ea <- nation_asia %>%
  filter(str_detect(Region, "Eastern Asia")) %>%
  select(-(Region))


asia_ea_import <- ggplot(nation_asia_ea,
       aes(x = `Month-Year`, y = Amount.x, color = Country)) +
  geom_line()+
  labs (x = "Date",
        y = "Import in Thousand Dollars",
        fill = "Country",
        title = "Import in Thousand Dollars by east Asian Country") +
  scale_x_date(date_labels = "%b-%y")+
  theme_bw() 

asia_ea_export <- ggplot(nation_asia_ea,
       aes(x = `Month-Year`, y = Amount.y, color = Country)) +
  geom_line()+
  labs (x = "Date",
        y = "Export in Thousand Dollars",
        fill = "Country",
        title = "Export in Thousand Dollars by East Asian Country") +
  scale_x_date(date_labels = "%b-%y")+
  theme_bw() 

patchwork <- (asia_ea_import / asia_ea_export)
patchwork
```

```{r}
#| fig-height: 12
#| fig-width: 14
#| code-fold: true

ggplot(nation_asia_ea,
  aes(x = `Month-Year`, y = Amount.x, color = Country)) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Date",
       y = "Net Trade",
       fill = "Country",
       title = "Net Trade by East Asian Country") +
  theme(legend.position = "top") +
  theme_bw() +
  geom_point() +
  transition_reveal(nation_asia_ea$`Month-Year`)
```

#### Insights on the East Asian Region

**Singapore has the highest Import and Export Trade relationship with Mainland China.** This is due to being supported by the updated bilateral agreement made in 2019. Given Singapore is one of the biggest trade hubs and China as a manufacturing hub, Merchandise Trade between these nations are part of the supply chain such as electronic equipment, machinery and minerals.

**Singapore - China Net Trade is higher in favor of Singapore (higher export).** This is due to Singapore exporting more variety of goods from consumer to manufacturing requirements to China.

::: callout-note
##### Singapore - China Trade Relationship

**SINGAPORE-CHINA**

In 2020, [Singapore](https://oec.world/en/profile/country/sgp) exported \$42.9B to [China](https://oec.world/en/profile/country/chn). The main products that [Singapore](https://oec.world/en/profile/country/sgp) exported to [China](https://oec.world/en/profile/country/chn) are [Integrated Circuits](https://oec.world/en/profile/hs/integrated-circuits) (\$10.6B), [Ethylene Polymers](https://oec.world/en/profile/hs/ethylene-polymers) (\$2.25B), and [Machinery Having Individual Functions](https://oec.world/en/profile/hs/machinery-having-individual-functions) (\$1.86B).

**CHINA-SINGAPORE**

In 2020, [China](https://oec.world/en/profile/country/chn) exported \$54B to [Singapore](https://oec.world/en/profile/country/sgp). The main products that [China](https://oec.world/en/profile/country/chn) exported to [Singapore](https://oec.world/en/profile/country/sgp) were [Integrated Circuits](https://oec.world/en/profile/hs/integrated-circuits) (\$5.55B), [Refined Petroleum](https://oec.world/en/profile/hs/refined-petroleum) (\$4.88B), and [Broadcasting Equipment](https://oec.world/en/profile/hs/broadcasting-equipment) (\$3.97B). South East Asian Import Export Visualization
:::

**Singapore also have a high export low import relationship with Hong Kong.** This is due to Hong Kong primarily acting as Financial Services Sector with minimal merchandise export on its own and rely on imports for consumer goods.

### South East Asian Import Export Visualization

This section shows the trade relation in South East Asia.

```{r}
#| fig-height: 16
#| fig-width: 14
#| code-fold: true

nation_asia_sea <- nation_asia %>%
  filter(str_detect(Region, "South-eastern Asia")) %>%
  select(-(Region))


asia_sea_import <- ggplot(nation_asia_sea,
       aes(x = `Month-Year`, y = Amount.x, color = Country)) +
  geom_line()+
  labs (x = "Date",
        y = "Import in Thousand Dollars",
        fill = "Country",
        title = "Import in Thousand Dollars by Southeast Asian Country") +
  scale_x_date(date_labels = "%b-%y")+
  theme_bw() 

asia_sea_export <- ggplot(nation_asia_sea,
       aes(x = `Month-Year`, y = Amount.y, color = Country)) +
  geom_line()+
  labs (x = "Date",
        y = "Export in Thousand Dollars",
        fill = "Country",
        title = "Export in Thousand Dollars by Southeast Asian Country") +
  scale_x_date(date_labels = "%b-%y")+
  theme_bw() 

patchwork <- (asia_sea_import / asia_sea_export)
patchwork
```

```{r}
#| fig-height: 12
#| fig-width: 14
#| code-fold: true

ggplot(nation_asia_sea,
  aes(x = `Month-Year`, y = Amount.x, color = Country)) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Date",
       y = "Net Trade",
       fill = "Country",
       title = "Net Trade by Southeast Asian Country") +
  theme(legend.position = "top") +
  theme_bw() +
  geom_point() +
  transition_reveal(nation_asia_sea$`Month-Year`)
```

#### Insights on the South East Asian Region

**Malaysia is the top import and export trade partner with Singapore.** This is due to proximity between the nations. Furthermore. Net Trade has higher export with Singapore

::: callout-note
##### Singapore - Malaysia Trade Relationship

**MALAYSIA-SINGAPORE**

In 2020, [Malaysia](https://oec.world/en/profile/country/mys) exported \$36.5B to [Singapore](https://oec.world/en/profile/country/sgp). The main products that [Malaysia](https://oec.world/en/profile/country/mys) exported to [Singapore](https://oec.world/en/profile/country/sgp) are [Integrated Circuits](https://oec.world/en/profile/hs/integrated-circuits) (\$11.1B), [Refined Petroleum](https://oec.world/en/profile/hs/refined-petroleum) (\$5.63B), and [Office Machine Parts](https://oec.world/en/profile/hs/office-machine-parts) (\$1.24B).

**SINGAPORE-MALAYSIA**

In 2020, [Singapore](https://oec.world/en/profile/country/sgp) exported \$22.9B to [Malaysia](https://oec.world/en/profile/country/mys). The main products that [Singapore](https://oec.world/en/profile/country/sgp) exported to [Malaysia](https://oec.world/en/profile/country/mys) were [Integrated Circuits](https://oec.world/en/profile/hs/integrated-circuits) (\$5.7B), [Refined Petroleum](https://oec.world/en/profile/hs/refined-petroleum) (\$4.33B), and [Gold](https://oec.world/en/profile/hs/gold) (\$894M).
:::

**Singapore has a high export to Indonesia due to the high amounts of Gold and manufacturing parts sent over.** Similar to nations that Singapore does bilateral trades of similar good, higher exports are primarily due to higher variety of goods.

## European Trade Analysis - Visual Exploration

This section explores the macro economic trade relation of Singapore to its European Trade partners.

### Europe Sub Region Visualization

```{r}
#| fig-height: 16
#| fig-width: 14
#| code-fold: true

ggplot(nation_eu, aes(x = Amount.x, y = Amount.y, size = Net_Trade, color = Country)) + 
  geom_point(show.legend = FALSE, alpha = 0.7) +
  scale_color_viridis_d() +
  scale_size(range = c(2, 12)) +
  transition_time(nation_eu$`Month-Year`) +
  labs(title = "Animated Scatterplot per European Region per Year: {frame_time}", x = "Imports", y = "Exports", fill = "Region") +
  theme(legend.position = "bottom") + 
  theme_bw() + 
  facet_wrap(~Region)
```

Based on the animated scatter plot, Western European nations have more dynamic trade relations as compared to others

### Western European Import Export Visualization

This section shows the trade relation in West Europe.

```{r}
#| fig-height: 16
#| fig-width: 14
#| code-fold: true

nation_eu_west <- nation_eu %>%
  filter(str_detect(Region, "Western Europe")) %>%
  select(-(Region))


eu_west_import <- ggplot(nation_eu_west,
       aes(x = `Month-Year`, y = Amount.x, color = Country)) +
  geom_line()+
  labs (x = "Date",
        y = "Import in Thousand Dollars",
        fill = "Country",
        title = "Import in Thousand Dollars by West Euro Country") +
  scale_x_date(date_labels = "%b-%y")+
  theme_bw() 

eu_west_export <- ggplot(nation_eu_west,
       aes(x = `Month-Year`, y = Amount.y, color = Country)) +
  geom_line()+
  labs (x = "Date",
        y = "Export in Thousand Dollars",
        fill = "Country",
        title = "Export in Thousand Dollars by West Euro Country") +
  scale_x_date(date_labels = "%b-%y")+
  theme_bw() 

patchwork <- (eu_west_import / eu_west_export)
patchwork
```

```{r}
#| fig-height: 12
#| fig-width: 14
#| code-fold: true

ggplot(nation_eu_west,
  aes(x = `Month-Year`, y = Amount.x, color = Country)) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Date",
       y = "Net Trade",
       fill = "Country",
       title = "Net Trade by West Euro Country") +
  theme(legend.position = "top") +
  theme_bw() +
  geom_point() +
  transition_reveal(nation_eu_west$`Month-Year`)
```

#### Insights on the Western European Region

As compared to Asian Trade partners, Western Europe experience high levels of fluctuation, presenting no clear front runner in trade relations with Singapore. **This may be due to West Europe being majority part of the European Union and its 'Single Market'.** The 'Single Market' allows companies in Europe to operate in all nation states and be able to cooperatively export their products as a block while import products outside in bulk as well.
