---
title: "Data Visualization Makeover 04"
subtitle: "Putting Visual Analytics into Practical Use"
author: "Bryant Philippe Lee"
date: February 26, 2022
---

## Objective

The objective is to uncover the impact of COVID-19 as well as the global economic and political dynamic in 2022 on Singapore bi-lateral trade (i.e. Import, Export and Trade Balance) by using appropriate analytical visualisation techniques to enhance user and data discovery experiences.

The visualization is created using the **Merchandise Trade** provided by **Department of Statistics, Singapore (DOS)*.***\

## Getting Started - Data Loading and Processing

### Installing and loading the required libraries

```{r}
pacman::p_load(scales, viridis, lubridate, ggthemes, gridExtra, tidyverse, readxl, knitr, data.table, readxl)
```

### Importing Data

This code chunk is to import the data from Merchandise Trade Excel File. The following sheets are within the file:

-   T1 = Import Data

-   T2 = Export Data

The data shows the import export relationship between Singapore and the other nations in terms of Million Dollar or Thousand Dollar

```{r}
import_data <- read_excel("data/Merchandise_Trade.xlsx", sheet = "T1")
import_data
```

```{r}
export_data <- read_excel("data/Merchandise_Trade.xlsx", sheet = "T2")
export_data
```

::: callout-note
It is observed that the dataset is presented by the following:

-   columns are listed per month-year, as such must be converted to date time format to ensure order

-   each nation (row) is listed either Thousand Dollars or Million Dollars, as such dataset must be corrected to ensure scale
:::
