---
title: "Data Visualization Makeover 03"
subtitle: "Putting Visual Analytics into Practical Use"
author: "Bryant Philippe Lee"
date: February 15, 2022
---

## Objective

The objective of this exercise is to uncover the salient patterns of the resale prices of public housing property by residential towns and estates in Singapore by using appropriate analytical visualisation techniques and apply appropriate interactive techniques to enhance user and data discovery experiences.

The visualization is created using the ***Resale flat princes based on registration date from Jan-2017 onwards ([Data.gov.sg](https://data.gov.sg/)).***

### Scope and Limitation

For the purpose of this study, the focus should be on **3-ROOM, 4-ROOM and 5-ROOM types**. You can choose to focus on either one housing type or multiple housing types. The study period should be on **2022**.

## Getting Started - Data Loading and Processing

### Installing and loading the required libraries

Before we get started, it is important for us to ensure that the required R packages have been installed.

```{r}
pacman::p_load(tidyverse, ggplot2, knitr, plotly, skimr, questionr, funModeling, sf, tmap, quanteda, NLP, ggiraph, ggstatsplot, dplyr, crosstalk, DT, ggdist, gganimate, ggpubr, corrplot, performance, parameters, see)
```

### Importing Data

This code chunk is to import the data from resale-flat-prices-based-on-registration-date-from-jan-2017-onwards.csv file to the Quarto/R page.

```{r}
total_flat_data <- read_csv("data/resale-flat-prices-based-on-registration-date-from-jan-2017-onwards.csv")
```

### Filtering the Dataset

This section deals with limiting the data to the following criteria:

-   The year should be **2022**

-   The flat_type should be only **3-ROOM, 4-ROOM and 5-ROOM**

::: callout-note
This code uses the dplyr package to filter out the rows with year 2022 and correct flat type
:::

```{r}
filter_flat_data <- filter(total_flat_data, grepl('2022', month) & flat_type %in% c("3 ROOM", "4 ROOM","5 ROOM"))

filter_flat_data
```

### Exploring the Data

This section of the code aims to ensure that the scope and limitation are followed and the data are listed correctly

```{r}
skimr::skim(filter_flat_data)
```

```{r}
glimpse(filter_flat_data, 60)
```

```{r}
unique(filter_flat_data$month)
```

```{r}
unique(filter_flat_data$flat_type)
```

```{r}
unique(filter_flat_data$storey_range)
```

::: callout-note
##### Critical!

storey_range is listed as a string and needed for the arrangement to be corrected moving forward
:::

```{r}
unique(filter_flat_data$floor_area_sqm)
```

```{r}
unique(filter_flat_data$remaining_lease)
```

::: callout-note
##### Critical!

remaining lease is listed as a string and needed for the arrangement to be corrected moving foreward
:::

```{r}
summary(filter_flat_data$resale_price)
```

### Correcting the Data

This section deals with correcting the data last inspected previously.

#### Correcting the Storey Arrangement

The code adds corrects the order of the storey arrangement

```{r}
storey_correct <- c("01 TO 03", "04 TO 06", "07 TO 09", "10 TO 12", "13 TO 15", "16 TO 18", "19 TO 21", "22 TO 24", "25 TO 27", "28 TO 30", "31 TO 33", "34 TO 36", "37 TO 39", "40 TO 42", "43 TO 45", "46 TO 48", "49 TO 51")  

use_flat_data <- filter_flat_data %>%
  mutate (storey_range = factor(storey_range, levels = storey_correct)) %>%
  ungroup()
```

```{r}
ggplot(data = use_flat_data,
       aes(y = storey_range)) +
  geom_bar() 
```

#### Converting Remaining Lease from String to Num

The code adds a new integer column derived from the remaining_lease string column.

::: callout-note
This code uses the gsub-regex function to get the integer within the string.
:::

```{r}
lease_ryear <- as.numeric(gsub("([0-9]+).*$", "\\1", filter_flat_data$remaining_lease))

use_flat_data$lease_ryear <- lease_ryear

use_flat_data
```

## Data Exploration through Statistical Visualization

This section presents the statistical visualization and explanation of the data.

### Analysis on the Type of Flat

This section presents insights with regards to the relationship between Type of Flat and other variables.

#### One-way ANOVA Test on Type of Flat and Resale Price

```{r}
#| fig-height: 12
#| fig-width: 12
#| code-fold: true

ggbetweenstats(
  data = use_flat_data,
  x = flat_type,
  y = resale_price, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

Coming from One-way ANOVA test between Type of Flat and the Resale Price, overall p-value of the ANOVA test is 0.00 \< 0.05 means that we can reject the null hypotheses meaning each flat type have different mean resale price coming from the ANOVA hypothesis:

-   Ho = there's no difference between the means and conclude that a significant difference does not exist.

-   Ha = there's difference between the means and conclude that a significant difference does exist.

Pairwise P-Test also concluded the following:

|    Relation     |      P-Value      |                     Remark                     |
|:-----------------:|:-----------------:|:--------------------------------:|
| 3 Room - 4 Room | 3.25 e-08 \< 0.05 | Reject null hypothesis, Significant difference |
| 3 Room - 5 Room |     0 \< 0.05     | Reject null hypothesis, Significant difference |
| 4 Room - 5 Room |     0 \< 0.05     | Reject null hypothesis, Significant difference |

::: callout-note
##### P-Value = 0

Please note that P-Value cannot be equals to 0, the 0 value is a reference that p-value is very close to 0
:::

It can also be noted that in 2022, around 46% (n = 11,312 rooms) of the total scoped rooms belong to the 4 Room category and with 54% of the rooms split almost evenly between 3 and 5 Room categories.

#### One-way ANOVA Test on Type of Flat and Floor Sqr Area

```{r}
#| fig-height: 12
#| fig-width: 12
#| code-fold: true

ggbetweenstats(
  data = use_flat_data,
  x = flat_type,
  y = floor_area_sqm, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

Coming from One-way ANOVA test between Type of Flat and the Floor Area, overall p-value of the ANOVA test is 0.00 \< 0.05 means that we can reject the null hypotheses meaning each flat type have different mean resale price coming from the ANOVA hypothesis:

-   Ho = there's no difference between the means and conclude that a significant difference does not exist.

-   Ha = there's difference between the means and conclude that a significant difference does exist.

Pairwise P-Test also concluded the following:

|    Relation     |      P-Value      |                     Remark                     |
|:-----------------:|:-----------------:|:--------------------------------:|
| 3 Room - 4 Room | 1.19 e-08 \< 0.05 | Reject null hypothesis, Significant difference |
| 3 Room - 5 Room |     0 \< 0.05     | Reject null hypothesis, Significant difference |
| 4 Room - 5 Room | 1.07 e-09 \< 0.05 | Reject null hypothesis, Significant difference |

#### Correlation between Resale Price, Remaining Lease Years and Floor Area

```{r}
#| code-fold: true

ggstatsplot::ggcorrmat(
  data = use_flat_data,
  cor.vars = c(resale_price, lease_ryear, floor_area_sqm)
)
```

Based on the correlation coefficients, there is moderate correlation between resale price and floor area and resale price and remaining lease year while weak correlation between the floor area and remaining lease year. This concludes that there is a relationship between resale price and remaining lease year and resale price and floor area though this does not mean causality between both relationship.

#### Model Diagnostic 

```{r}
#| fig-height: 36
#| fig-width: 36
#| code-fold: true

model <- lm(resale_price ~ flat_type + lease_ryear + floor_area_sqm, data = use_flat_data)

check_model(model)
```

### Analysis on 3 Room Flats

```{r}
room_3 <- filter(use_flat_data, grepl("3 ROOM", flat_type))
room_3
```

#### One-way ANOVA Test on Town and Resale Price

```{r}
#| fig-height: 12
#| fig-width: 12
#| code-fold: true

ggbetweenstats(
  data = room_3,
  x = town,
  y = resale_price, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

```{r}
res.aov <- aov(resale_price~town, data = room_3)
summary (res.aov)

TukeyHSD(res.aov)
```

::: callout-note
##### Too Many Values

There is an estimated 55 areas as such would yield large amount of combination for pairwise analysis.
:::

#### One-way ANOVA Test on Flat Model and Resale Price

```{r}
#| fig-height: 12
#| fig-width: 12
#| code-fold: true

ggbetweenstats(
  data = room_3,
  x = flat_model,
  y = resale_price, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

Based on the Oneway ANOVA, overall ANOVA has a p-value less than 0.05 (4.36 e-107 \< 0.05), indicating that the mean price is significantly different from each Flat Model.

#### Correlation between Resale Price, Remaining Lease Years and Floor Area

```{r}
#| code-fold: true

ggstatsplot::ggcorrmat(
  data = room_3,
  cor.vars = c(resale_price, lease_ryear, floor_area_sqm)
)
```

Based on the correlation coefficients, there is moderate correlation between resale price and remaining lease year while weak to no correlation between the floor area and remaining lease year and resale price and floor area. This concludes that there is a relationship between resale price and remaining lease year though this does not mean causality between both relationship.

### Analysis on 4 Room Flats

```{r}
room_4 <- filter(use_flat_data, grepl("4 ROOM", flat_type))
room_4
```

#### One-way ANOVA Test on Flat Model and Resale Price

```{r}
#| fig-height: 12
#| fig-width: 12
#| code-fold: true

ggbetweenstats(
  data = room_4,
  x = flat_model,
  y = resale_price, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

Based on the Oneway ANOVA, overall ANOVA has a p-value less than 0.05 (5.37 e-34 \< 0.05), indicating that the mean price is significantly different from each Flat Model.

#### Correlation between Resale Price, Remaining Lease Years and Floor Area

```{r}
#| code-fold: true

ggstatsplot::ggcorrmat(
  data = room_4,
  cor.vars = c(resale_price, lease_ryear, floor_area_sqm)
)
```

Based on the correlation coefficients, there is moderate correlation between resale price and remaining lease year while weak to no correlation between the floor area and remaining lease year and resale price and floor area. This concludes that there is a relationship between resale price and remaining lease year though this does not mean causality between both relationship.

### Analysis on 5 Room Flats

```{r}
room_5 <- filter(use_flat_data, grepl("5 ROOM", flat_type))
room_5
```

#### One-way ANOVA Test on Flat Model and Resale Price

```{r}
#| fig-height: 12
#| fig-width: 12
#| code-fold: true

ggbetweenstats(
  data = room_5,
  x = flat_model,
  y = resale_price, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

Based on the Oneway ANOVA, overall ANOVA has a p-value less than 0.05 (3.88 e-17 \< 0.05), indicating that the mean price is significantly different from each Flat Model.

#### Correlation between Resale Price, Remaining Lease Years and Floor Area

```{r}
#| code-fold: true

ggstatsplot::ggcorrmat(
  data = room_5,
  cor.vars = c(resale_price, lease_ryear, floor_area_sqm)
)
```

Based on the correlation coefficients, for 5 Room Flats, the resale price, remaining lease years and floor area have weak to no correlation meaning each variable is independent of each other.

## Learning and Insights 

**Size Limitations:** due to the limitations of quarto size, visualization many variables may be difficult. Filters and selected values may be considered

**Ease of Statistical Understanding:** certain visualizations may not be easily understood by readers/users. Simplification may be a better option.
